A Simple Model of Energy Transition Under a Carbon Budget

### 2015-05-30 Final_Ciccone_MS_Thesis.tex
Final draft attempt 2

%\documentclass[singlespace]{easychithesis}

\documentclass{easychithesis}

@@

\title{A Simple Model of Energy Transition Under a Carbon Budget} 
\author{Adriana Ciccone}
\date{\today}
\department{Environmental Science \& Policy}
\division{The Harris School of Public Policy Studies} 
\degree{Master of Science} 
\maketitle

@@

\tableofcontents

\mainmatter

\chapter{Introduction}
\section{Motivation}

\paragraph{} While there is emerging political consensus in the United States that climate change is occurring and will have significant economic and social impacts, there remain significant barriers to a low-carbon economy. Electric power generation accounts for roughly 30\% of total greenhouse gas emissions in the United States, the largest of any sector \cite{ElectricGHGEmit}. While there now exist renewable power generation sources, such as wind turbines and photovoltaic solar panels, costs remain high compared to traditional coal and natural gas power plants, though there is reason to believe that costs will decrease over time as technology improves and economies of scale begin to take effect. With scientific consensus to limit further greenhouse gas emissions to limit warming to an acceptable level, the problem that results is that of energy transition from $CO_2$ intensive to renewable technologies.

\paragraph{} Reducing emissions can therefore be accomplished by replacing durable fossil fuel infrastructure. Existing infrastructure, however, may still have usable life remaining, and discarding it by early retirement is wasteful considering that there would need to be additional investment in new capital to meet total energy demands. However, in the event of an international treaty or other domestic regulation limiting greenhouse gas emissions, such early retirement may become necessary. In this paper i propose a simple model of a low-cost transition to clean power generation with exactly these constraints; a fixed energy demand and a carbon budget.

\paragraph{} The model proposed on the following pages seeks to answer the question of how best to transition to clean power generation. This model is intentionally simplified to several key equations to highlight the relationship between key parameters and the outcome. It is not intended to be an exact estimate of the optimal profile of clean energy investments nor give an exact value of the amount of capital that needs to be retired early. Instead I write down the simplest possible model that will allow us to develop an intuition for the key drivers of capital accrual and early retirement.

\paragraph{} There are two sectors that produce electricity in this model, high-emitting or dirty capital such as coal plants and natural gas generators. The other sector is low-emitting or clean generating capital such as wind turbines and photovoltaic solar panels. Wind and solar are more expensive to build than gas or coal plants for the same unit of energy produced. However, neither wind nor solar emit the amount per kilowatt-hour of carbon dioxide that the dirty plants do, and thus will not be subjected to any international or domestic regulation regarding greenhouse gas emissions.

\paragraph{} The optimization requires minimizing the cost of building new capital in both sectors while meeting the domestic energy demand and complying with an emissions cap that is determined exogenously. With these constraints, we modify key parameter to understand what factors impact the speed of transition and timing of any necessary early retirement of polluting plants. In addition, existing capital depreciates exponentially regardless of early retirement. This creates an opportunity to substitute clean power generation without having to early retire any coal or natural gas plants. Therefore we could gradually replace high-emitting capital with clean as the high-emitting stock depreciates naturally. However in certain cases, the rate of natural depreciation may not be sufficient to avoid exceeding the emissions cap. In these cases substitution of fully depreciated capital with clean capital would be required, but early retirement of high-emitting capital would also be required. Additionally, the rate at which the cost of clean capital decreases may affect the optimal investment profile, if it becomes feasible that clean energy capital can be cost competitive with existing coal and gas plants. To this effect, I develop this model and run simulations with a range of parameters to show the sensitivity of the optimal transition cost to each of these parameters.

\section{Prior Work}

@@

\chapter{Modeling Choices}
\section{Model Description}\label{sec:ModelDesc}

\paragraph{} The region or country of interest produces electricity by two methods: clean (low emitting, represented by $k^l(t)$) and dirty (high emitting, represented by $k^h(t)$). Clean capital produces electricity with low carbon dioxide emissions, but, as a newer technology, it is more expensive per unit of power output than dirty. Since it is a new technology, the capital cost is expected to decrease over time as the production process matures and technological innovation produces lower-cost units. Dirty capital is a mature technology that is cheap to build but emits a higher level of carbon dioxide per unit of electricity produced than clean. Since it is a mature technology, there is little reason to believe that production costs will decline substantially in the future.

\paragraph{} Each type of capital depreciates exponentially over time. The decision variables for the optimization are additional investments in each type of capital that also depreciate at the same rate, represented by four functions ($H^+(t), H^-(t), L^+(t), \& L^-(t)$). These decision variables define the optimal path of investment in dirty (high emitting) and clean (low emitting) capital. All four functions are non-negative over their domain. The functions $H^+(t)$ and $L^+(t)$ represent additional dollars of investment to build new capital of each type. $H^-(t)$ and $L^-(t)$ remove capital from the economy via early retirement. There is no scrap cost to early retirement - the economy simply loses the ability to produce electricity from that particular facility. As I show in the next section, this has consequences for the ability of the economy to meet it's total electricity demand. The retired plant cannot be re-started. Any need for additional capital in time t must be met by a positive investment in either $H^+(t)$ or $L^+(t)$.

\begin{equation}\label{eq:simpleHCapitalConstraint}
k^h(t) = H_0 e^{-t/n} + \int_0^t e^{x-t/n} H^+(x) dx - \int_0^t H^-(x)dx \geq 0 \qquad \forall t
\end{equation}

\begin{equation}\label{eq:simpleLCapitalConstraint}
k^l(t) = L_0 e^{-t/n} + \int_0^t e^{x-t/n} L^+(x) dx - \int_0^t L^-(x)dx \geq 0 \qquad \forall t
\end{equation}

These equations express total capital in each time period as a function of some exponential depreciation and all prior investments. At each time t, the initial capital endowment, $H_0 = H^+(0)$ and $L_0 = L^+(0)$ depreciate exponentially and any further positive investments depreciate based on their age (x-t). The integral in the above expression shows how each investment $H^+(t)$ or $L^+(t)$ is weighted by an exponential term of the difference in that investment's age and the current time period. Therefore we see that early investments have a more negative exponential term and count less towards existing capital than more recent investments. While we have no limits on the amount of capital in each time period other than the non-negativity constraint on the decision variables, a physical limitation remains that we cannot have negative amounts of capital. We therefore restrict both $k^h(t)$ and $k^l(t)$ to be non-negative in all time periods.

@@

\paragraph{} Note that in this objective function, there is no benefit to early retirement in the form of reduced future operating costs. In the case of dirty technologies, investing in new capital involves pre-purchasing all the required coal or natural gas, as well as labor, operating and maintenance costs. When a plant is retired early, these variable operating costs don't get redeemed, which biases the model towards investing in clean technologies earlier by penalizing early retirement more than a real-world scenario would. We address this problem in the extended model in section \ref{VintExplan}.

\paragraph{} The economy has some total energy demand that can follow any exogenous path over time. This represents total annual electricity demand, in billions of kilowatt-hours, and therefore daily and seasonal fluctuations in power demand are ignored. Capital must be used to produce electricity to meet this demand, and we must convert from dollars of capital to kilowatt-hours of electricity to take in to account the relative cost difference of producing one unit of clean energy versus one unit of dirty energy.

@@

In this equation, $F^h (t)$ and $F^l (t)$ are the overnight capital cost per billion kilowatt-hours per year. That is, they represent the number of kilowatt-hours per year that each dollar of capital can produce. In early years when the clean technology is still relatively new and expensive, a dollar of capital may not produce much energy per year, but a dollar of clean capital invested later in the model may produce significantly more energy. It is important to be clear however, on what this cost is. It is not an annual maintenance cost or any measure of financial obligation for additional production. It is simply the the ability of that unit (dollar) of capital to produce a certain amount of electricity.

\paragraph{} An emissions cap is negotiated by the region's government. Over the course of a fixed number of years, the region must not emit more than a negotiated percent of business as usual emissions. In practice, the government could negotiate to agree to a limit on total number of tons of $CO_2$ emitted instead of negotiating a percentage of business as usual. However, it is convenient to represent a bulk emissions cap as a percent reduction for both modeling and to understand how our optimal cost will change as a function of total allowed emissions.

@@

\begin{equation}\label{eq:dLdL+}
\frac{d\Lagr}{dH^+(t)} = 1/r(e^{-rN} -1) + n(1-e^{-t/n})(\pi - \lambda F^l(1-e^{-t/\tau})) + \gamma = 0
\end{equation}

\begin{equation}\label{eq:dLdL-}
\frac{d\Lagr}{dH^-(t)} = (\lambda F^l(1-e^{-t/\tau}) - \pi)t + \delta = 0
\end{equation}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Results}
\section{Numerical Simulation}

@@

\paragraph{} Energy demand is exogenous and comes directly from the EIA. From the EIA demand forecast, I subtract generating capacity supplied by hydroelectric dams and nuclear plants to evaluate only the subset of demand that will be fulfilled by the clean (wind, solar photovoltaic) and dirty (coal, natural gas) technologies. I consider this an acceptable simplification that can still show the dynamics of the clean energy transition, though future work may want to include hydro and nuclear as well as separate wind, solar, gas and coal in to separate sectors. Energy demand growth is roughly linear over time, going from 2836 billion kWh in 2011 to 3734 billion kWh in 2040. For any simulation length greater than 30 years, I assume that growth will continue to grow at the same per year rate.

@@

\paragraph{} Of course, the main source of uncertainty in this model is the speed of cost reductions for clean energy capital. I model this forecast deterministically as a logistic function, with some initial low value and a maximum number of kWh/year per dollar that represents the cost floor of a mature technology. Varying the speed of transition between the minimum and maximum kWh/year/dollar allows us to understand the sensitivity of the objective function with respect to the speed by which clean capital cost decreases. Figure \ref{fig:FlScaleScenarios} gives some cost curve scenarios in comparison with the linear high-emitting capital cost curve. The minimum cost in each of these figures is half of the current weighted average overnight capital cost from the EIA and the maximum is the cost per kWh/year of natural gas. Both the minimum and the maximum are then adjusted to account for a lower renewable capacity factor (30\%) that accounts for the intermittent nature of renewable power generation. Emissions per kWh of clean energy are assumed to be 0; that is, any $C0_2$ emissions from the construction of the wind turbines or photovoltaic cells are assumed to be negligible.

@@

\section{Model Results}
\FloatBarrier

@@

\begin{figure}[h]
\includegraphics[width=\textwidth]{cap_and_invest_results_A_75.png}
\caption{Alpha = 75\label{fig:SimpleResultsA=75}}
\end{figure}

@@

\paragraph{} In comparison, when the emissions cap is set much lower, say 25\% of business as usual emissions, investment in clean technologies must happen much sooner and with a higher magnitude of investment. Such a transition is shown in figure \ref{fig:SimpleResultsA=25}.

\begin{figure}[h]
\includegraphics[width=\textwidth]{cap_and_invest_results_A_25.png}
\caption{Alpha = 25\label{fig:SimpleResultsA=25}}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[width=\textwidth]{cap_and_invest_results_Fl_1.png}
\caption{Fl = 1\label{fig:SimpleResultsFl=1}}
\end{figure}

@@

\paragraph{} Similar patterns in our optimization solution emerge when we consider the speed of capital cost reductions of clean technologies. A low speed transition, such as one shown in figure \ref{fig:SimpleResultsFl=1} shows that additional positive investments in high-emitting technologies must be made early in the simulation to ensure adequate generating capacity. However, much of this early investment must be retired early to avoid exceeding the emissions cap resulting in a higher penalty to the overall transition cost as less cost-effective clean capital is substituted. This is contrasted with a very quick cost reduction scenario such as in figure \ref{fig:SimpleResultsFl=100} where hardly any additional high-emitting capital is invested at all. Instead the model can meet it's generation demand by substituting clean energy capacity at a relatively lower cost per kilowatt hour than in the previous scenario. While some high emitting capital must still be retired early, the amount of capital subjected to early retirement is roughly an order of magnitude less than in the slow-transition speed scenario.

\begin{figure}[h]
\includegraphics[width=\textwidth]{cap_and_invest_results_Fl_100.png}
\caption{Fl = 100\label{fig:SimpleResultsFl=100}}
\end{figure}

@@

\section{Vintaged Model}\label{VintExplan}

@@

\chapter{Model Discussion}

@@

\singlespacing
\pagebreak
\addcontentsline{toc}{chapter}{References}

@@

\bibitem{EIATable5}
fill in

\bibitem{EIACapCost}
http://www.eia.gov/forecasts/capitalcost/pdf/updated\_capcost.pdf

\bibitem{ElectricGHGEmit}
http://www.epa.gov/climatechange/ghgemissions/sources/electricity.html

>>>>

\documentclass{article}

@@

\title{A Simple Model of Energy Transition Under a Carbon Budget} 
\author{Adriana Ciccone}
\date{\today}

\maketitle

@@

\section{Introduction}
\subsection{Motivation}

\paragraph{} While there is emerging political consensus in the United States that climate change is occurring and will have significant economic and social impacts, there remain significant barriers to a low-carbon economy. Electric power generation accounts for roughly 30\% of total greenhouse gas emissions in the United States, the largest of any sector \cite{ElectricGHGEmit}. While there now exist renewable power generation sources, such as wind turbines and photovoltaic solar panels, costs remain high compared to traditional coal and natural gas power plants. However, there is some reason to believe that costs will decrease over time as technology improves and economies of scale begin to take effect. With scientific consensus to limit further greenhouse gas emissions to limit warming to an acceptable level, the problem that results is that of energy transition from $CO_2$ intensive to renewable technologies.

\paragraph{} Reducing emissions can be accomplished by replacing durable fossil fuel infrastructure. Existing infrastructure, however, may still have usable life remaining, and discarding it by early retirement is wasteful, considering that there would need to be additional investment in new capital to meet total energy demands. However, in the event of an international treaty or other domestic regulation limiting greenhouse gas emissions, such early retirement may become necessary. In this paper i propose a simple model of a low-cost transition to clean power generation with exactly these constraints; a fixed energy demand and a carbon budget.

\paragraph{} The model proposed on the following pages seeks to answer the question of how best to transition to clean power generation. This model is intentionally simplified to the most basic of descriptive equations to highlight the relationship between key parameters and the outcome. It is not intended to be an exact estimate of the optimal profile of clean energy investments, nor give an exact value of the amount of capital that needs to be retired early. Instead I write down the simplest possible model that will allow us to develop an intuition for the key drivers of capital accrual and early retirement.

\paragraph{} There are two sectors that produce electricity in this model, high-emitting or dirty capital such as coal plants and natural gas generators. The other sector is low-emitting or clean generating capital such as wind turbines and photovoltaic solar panels. Wind and solar are more expensive to build than gas or coal plants for the same unit of energy produced. However, neither wind nor solar emit the  per-kilowatt-hour amount of carbon dioxide that the dirty plants do, and thus will not be subjected to any international or domestic regulation limiting greenhouse gas emissions.

\paragraph{} The optimization requires minimizing the cost of building new capital in both sectors while meeting the domestic energy demand and complying with an emissions cap that is determined exogenously. In addition, existing capital depreciates exponentially regardless of early retirement. This creates an opportunity to substitute clean power generation without having to early retire any coal or natural gas plants. Therefore we could gradually replace high-emitting capital with clean as the high-emitting stock depreciates naturally. However in certain cases, the rate of natural depreciation may not be sufficient to avoid exceeding the emissions cap. In these cases substitution of fully depreciated capital with clean capital would be required, but early retirement of high-emitting capital would also be required. Additionally, the rate at which the cost of clean capital decreases may affect the optimal investment profile, if it becomes feasible that clean energy capital can be cost competitive with existing coal and gas plants. To this effect, I develop this model and run simulations with a range of parameters to show the sensitivity of the optimal transition cost to each of these parameters.

\subsection{Prior Work}

@@

\section{Modeling Choices}
\subsection{Model Description}\label{sec:ModelDesc}

\paragraph{} The region or country of interest uses generating capital to produce electricity by two methods: clean (low emitting, represented by $k^l(t)$) and dirty (high emitting, represented by $k^h(t)$). Clean capital produces electricity with low carbon dioxide emissions, but, as a newer technology, it is more expensive per unit of power output than conventional fossil fuel-based generation. Since it is a new technology, the capital cost per kilowatt is expected to decrease over time as manufacturing processes mature and technological innovation produces more efficient designs. High emitting capital is a mature technology that is cheap to build but emits a higher level of carbon dioxide per unit of electricity produced than clean. Since it is a mature technology, there is little reason to believe that production costs will decline substantially in the future.

\paragraph{} Each type of capital depreciates exponentially over time. The decision variables for the optimization are additional investments or disinvestments in each type of capital, represented by four functions ($H^+(t)$, $H^-(t)$, $L^+(t)$, \& $L^-(t)$). These decision variables define the optimal path of investment in dirty (high emitting) and clean (low emitting) capital. All four functions are non-negative over their domain. The functions $H^+(t)$ and $L^+(t)$ represent additional dollars of investment to build new capital of each type. $H^-(t)$ and $L^-(t)$ remove capital from the economy via early retirement. There is no scrap cost to early retirement - the economy simply loses the ability to produce electricity from that particular facility. As I show in the next section, this has consequences for the ability of the economy to meet it's total electricity demand. The retired plant cannot be re-started. Any need for additional generating capital in time t must be met by a positive investment in either $H^+(t)$ or $L^+(t)$.

\begin{equation}\label{eq:simpleHCapitalConstraint}
k^h(t) = H_0 e^{-t/n^h} + \int_0^t e^{x-t/n^h} H^+(x) dx - \int_0^t H^-(x)dx \geq 0 \qquad \forall t
\end{equation}

\begin{equation}\label{eq:simpleLCapitalConstraint}
k^l(t) = L_0 e^{-t/n^l} + \int_0^t e^{x-t/n^l} L^+(x) dx - \int_0^t L^-(x)dx \geq 0 \qquad \forall t
\end{equation}

These equations express total capital in each time period as a function of some exponential depreciation and all prior investments. At each time t, the initial capital endowment, $H_0 = H^+(0)$ and $L_0 = L^+(0)$ depreciate exponentially and any further positive investments depreciate based on their age (x-t). The integral in the above expression shows how each investment $H^+(t)$ or $L^+(t)$ is weighted by an exponential term of the difference in that investment's age and the current time period. Therefore we see that early investments have a more negative exponential term and count less towards existing capital than more recent investments. The rate of depreciation is scaled by $n^h$ and $n^l$, which allows us to vary the lifetime of each type of capital. While we have no limits on the amount of capital in each time period other than the non-negativity constraint on the decision variables, a physical limitation remains that we cannot have negative amounts of capital. We therefore restrict both $k^h(t)$ and $k^l(t)$ to be non-negative in all time periods.

@@

\paragraph{} Note that in this objective function, there is no benefit to early retirement in the form of reduced future operating costs. In the case of dirty technologies, investing in new capital involves pre-purchasing all the required coal or natural gas, as well as labor, operating and maintenance costs. When a plant is retired early, any remaining fuel or variable operating costs don't get redeemed, biasing the model towards investing in clean technologies earlier by penalizing early retirement more than a real-world scenario would. We address this problem in the extended model in section \ref{VintExplan}.

\paragraph{} The economy has some total energy demand that can follow any exogenous path over time. This represents total annual electricity demand, in billions of kilowatt-hours, and therefore daily and seasonal fluctuations in power demand are ignored. Capital must be used to produce electricity to meet this demand, and we must convert from dollars of capital to kilowatt-hours of electricity produced per year to take in to account the relative cost difference of producing one unit of clean energy versus one unit of dirty energy.

@@

In this equation, $F^h (t)$ and $F^l (t)$ are the overnight capital cost per billion kilowatt-hours per year, plus the present value of expected future operating costs. That is, they represent the number of kilowatt-hours per year that each dollar of capital can produce. In early years when the clean technology is still relatively new and expensive, a dollar of capital may not produce much energy per year, but a dollar of clean capital invested later in the model may produce significantly more energy. It is important to be clear however, on what this cost is. It is not an annual maintenance cost or any measure of financial obligation for additional production. It is simply the the ability of that unit (dollar) of capital to produce a certain amount of electricity.

\paragraph{} An emissions cap is negotiated by the region's government. Over the course of a fixed number of years, the region must not emit more than a negotiated percent of business as usual emissions. In practice, the government could negotiate to agree to a limit on total number of tons of $CO_2$ emitted instead of negotiating a percentage of business as usual. However, it is convenient to represent a bulk emissions cap as a percent ($\alpha$) reduction for both modeling and to understand how our optimal cost will change as a function of total allowed emissions.

@@

\begin{equation}\label{eq:dLdL+}
\frac{d\Lagr}{dL^+(t)} = 1/r(e^{-rN} -1) + n(1-e^{-t/n})(\pi - \lambda F^l(1-e^{-t/\tau})) + \gamma = 0
\end{equation}

\begin{equation}\label{eq:dLdL-}
\frac{d\Lagr}{dL^-(t)} = (\lambda F^l(1-e^{-t/\tau}) - \pi)t + \delta = 0
\end{equation}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

@@

\paragraph{} Energy demand is exogenous and comes directly from the EIA. From the EIA demand forecast, I subtract generating capacity supplied by hydroelectric dams and nuclear plants to evaluate only the subset of demand that will be fulfilled by the clean (wind, solar photovoltaic) and dirty (coal, natural gas) technologies. I consider this an acceptable simplification that can still show the dynamics of the clean energy transition, since hydroelectric dams, barring some small-scale technological innovation, have reached their maximum penetration in the United States, and the future of nuclear is unclear. Future work may want to include hydro and nuclear as well as separate wind, solar, gas and coal in to separate sectors. Energy demand growth is roughly linear over time, going from 2836 billion kWh in 2011 to 3734 billion kWh in 2040. For any simulation length greater than 30 years, I assume that growth will continue to grow at the same per year rate.

@@

\paragraph{} Of course, the main source of uncertainty in this model is the speed of cost reductions for clean energy capital. I model this forecast deterministically as a logistic function, with some initial low value and a maximum number of kWh/year per dollar that represents the cost floor of a mature technology. Varying the speed of transition between the minimum and maximum kWh/year/dollar, as well as maximum output/\$ value allows us to understand the sensitivity of the objective function with respect to the speed by which clean capital cost decreases. Figure \ref{fig:FlScaleScenarios} gives some cost curve scenarios in comparison with the linear high-emitting capital cost curve. The minimum cost in each of these figures is half of the current weighted average overnight capital cost from the EIA and the maximum is the cost per kWh/year of natural gas. Both the minimum and the maximum are then adjusted to account for a lower renewable capacity factor (30\%) that accounts for the intermittent nature of renewable power generation. Emissions per kWh of clean energy are assumed to be 0; that is, any $C0_2$ emissions from the construction of the wind turbines or photovoltaic cells are assumed to be negligible.

@@

\subsection{Model Results}
\FloatBarrier

@@

\begin{figure}[h]
\includegraphics[width=\textwidth]{simple_cap_and_invest_results_Fl_1_alpha_75_FLFrac_25.png}
\caption{A slow transition speed with a lenient transition speed and high cost of renewable capital\label{fig:SimpleResultsA=75}}
\end{figure}

@@

\paragraph{} In comparison, when the emissions cap is set much lower, say 50\% of business as usual emissions, investment in clean technologies must happen much sooner and with a higher magnitude of investment. Such a transition is shown in figure \ref{fig:SimpleResultsA=25}.

\begin{figure}[h]
\includegraphics[width=\textwidth]{simple_cap_and_invest_results_Fl_10_alpha_50_FLFrac_50.png}
\caption{Medium transition speed with 50\% emissions reduction and moderate cost of renewable capital\label{fig:SimpleResultsA=25}}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[width=\textwidth]{simple_cap_and_invest_results_Fl_15_alpha_50_FLFrac_50.png}
\caption{Fast transition speed with 50\% emissions reduction and moderate cost of renewable capital\label{fig:SimpleResultsFl=1}}
\end{figure}

@@

\paragraph{} Similar patterns in our optimization solution emerge when we consider the speed of capital cost reductions of clean technologies. A fast speed transition, such as one shown in figure \ref{fig:SimpleResultsFl=1} shows that additional positive investments in high-emitting technologies may be made later in the simulation to ensure adequate generating capacity. However, much of this early investment must be retired early to avoid exceeding the emissions cap resulting in a higher penalty to the overall transition cost as less cost-effective clean capital is substituted. This is contrasted with a very quick cost reduction scenario such as in figure \ref{fig:SimpleResultsA=25} where hardly any additional high-emitting capital is invested at all. Instead the model can meet it's generation demand by substituting clean energy capacity at a relatively lower cost per kilowatt hour than in the previous scenario. While some high emitting capital must still be retired early, the amount of capital subjected to early retirement is roughly an order of magnitude less than in the slow-transition speed scenario.

@@

\subsection{Vintaged Model}\label{VintExplan}

@@

\begin{figure}[h]
\includegraphics[width=\textwidth]{vint_cap_and_invest_results_Fl_1_alpha_075_FLFrac_025.png}
\caption{Slow transition speed with lenient emissions reduction and high cost of renewable capital\label{fig:VintResultsFl=1_Alpha75_Frac25}}
\end{figure}

\begin{figure}[h]
\includegraphics[width=\textwidth]{vint_cap_and_invest_results_Fl_10_alpha_05_FLFrac_05.png}
\caption{Medium transition speed with 50\% emissions reduction and moderate cost of renewable capital\label{fig:VintResultsFl=10_Alpha50_Frac50}}
\end{figure}

\begin{figure}[h]
\includegraphics[width=\textwidth]{vint_cap_and_invest_results_Fl_15_alpha_05_FLFrac_05.png}
\caption{Fast transition speed with 50\% emissions reduction and moderate cost of renewable capital\label{fig:VintResultsFl=15_Alpha50_Frac50}}
\end{figure}

\section{Model Discussion}

@@

\pagebreak
\addcontentsline{toc}{chapter}{References}

@@

\bibitem{EIATable5}
Energy Information Administration. Annual Energy Outlook 2014

\bibitem{EIACapCost}
Energy Information Administration. Updated Capital Cost Estimates for Utility Scale Electricity Generating Plants. April 2013

\bibitem{ElectricGHGEmit}
Environmental Protection Agency. Sources of Greenhouse Gas Emissions.  http://www.epa.gov/climatechange/ghgemissions/sources/electricity.html . Accessed 4/25/2015

http://www.eia.gov/forecasts/capitalcost/
